Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-20 ICML2019/main.tex
pass on section 3

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
	\item We explored two ways to improve the accuracy: 1) calibrating noise to the sensitivity of a metric over distributions (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) Using a smooth upper bound on the local sensitivity and scale the noise to this smooth bound rather than global sensitivity, to improve the mechanism accuracy.
  \item A full proof on the newly designed mechanism is $(\epsilon, \delta)-$differential privacy is given in paper.
	\item We implemented the new proposed mechanism and other art-of-state mechanisms, comparing the performance in terms of accuracy and privacy.
\end{itemize}

@@

\noindent \textbf{Problem setting.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in Section \ref{sec_background}.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta and Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a private version of the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
\todo{However, distance between
distributions are more naturally to be measured by distribution
metrics such as $f$-divergence, Hellinger distance, etc., than the $\ell_1$ norm shown above.}
{\color{blue} changed a bit, still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

We will use the \emph{Hellinger} distance $\hellinger(\cdot,\cdot)$.
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the square of the Hellinger distance between them is defined as:
\[
\hellinger(P,Q)^2 =\frac{1}{2} \int \bigg(\frac{dP}{d \lambda} - \frac{dQ}{d \lambda}\bigg)^2 d \lambda
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$ the above can be reduced to: 
\[
 \hellinger(P,Q) =  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
Our choice to use Hellinger distance is motivated by three facts:
\begin{itemize}
  \item It simplifies calculations in the case of the probabilistic models considered here.

@@

  \item The accuracy can be improved by using a smooth bound on Hellinger distance's local sensitivity. Indeed, as shown in Figure \ref{fig_sensitivity}, there can be a gap
    between the global sensitivity and the local sensitivity.
\begin{figure}[ht]
\centering
\includegraphics[width=0.32\textwidth]{sensitivity_2d.eps}
\caption{{The local sensitivity of $\hellinger$}}
\label{fig_sensitivity}
\end{figure}
\end{itemize}

\section{Private Mechanisms}
\label{sec_mechs}
\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
We start off by presenting a mechanism based on the Laplace primitive.
The mechanism, denoted by {\color{red}{i think this macro has too many arguments}}($\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$),
simply adds Laplace noise to the parameters of the posterior distributions. The mechanisms also performs a
post-processing phase by {\color{red}{flooring}} and  truncating the noisy parameters so that they belong to the right interval.
In the case of the Beta-Binomial model the output produced is: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$
with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into
$[0,n]$. This mechanism is described by the pesudocode in Figure \ref{mech:lapmech}.\\
Since we are adding noise directly to the parameter, the sensitivity should be calibrated to $\ell_1$ of two Beta distributions:
$
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
$,
i.e., proportional to the dimensionality.
  \begin{algorithm}
  \caption{$\lapmech$ in Beta-binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad  $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad  $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad  $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ \\
  % \RETURN $a_{k+1}$.
  \noindent {\bf return} $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{Improving the sensitivity ($\ilapmech$)}
 The accuracy of $\lapmech$ can be reduced to a constant.
 Indeed, two adjacent data sets $\dataobs, \dataobs'$  induce posterior distributions -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ --
 whose parameters differ by at most 2 w.r.t $\ell_1$ norm. This because we can see the parameters of the distributions as an histogram.
 Hence, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.
 {\color{red} Notice that we are using the bounded notion of differential privacy \cite{????}.}
 \todo{why 1 in 2 dimensions?}
 We denote the mechanism which calibrates the noise to this smaller sensitivity as $\ilapmech$.
 $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0) $
 where $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in the Beta-binomial model and
 $\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots,  \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}$
 with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Figure \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

\subsection{A new mechanism}
In this section, we explore a new mechanism which we denote by $\expmech$.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{Exponential Mechanism with Hellinger Distance Metric ($\expmech$)}
Our exploration starts from the standard exponential mechanism $\expmech(x,u,\candidateset)$, which samples an element from the candidate set $\candidateset$ with probability proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$:
\[
\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
{\exp(\frac{\epsilon u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon u(x,r')}{2 GS})},
\]
where $u(x,r)$ is the Hellinger scoring function over candidates, $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function (i.e. the global sensitivity of the Bayesian inference w.r.t the Hellinger distance), calculated by:
\begin{equation*}
GS = 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}

\subsubsection{Exponential Mechanism with Hellinger Metric and Local Sensitivity ($\lexpmech$)}
\label{subsec_emls}
In order to achieve good accuracy, we make an attempt on the Exponential mechanism with local sensitivity $\lexpmech(x,u,\candidateset)$. This is just an attempt since local sensitivity is non-private. $\lexpmech$ shares the same candidate set and utility function as $\expmech$, outputting a result $r \in \mathcal{R}$ with probability proportional to $\exp(\frac{\epsilon u(x,r)}{2 LS(\dataobs)})$:
\[
\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
{\exp(\frac{\epsilon u(x,r)}{2 LS(\dataobs)})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon u(x,r')}{2 LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of scoring function, calculated by:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)=\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert.
\end{equation}

Thanks to the symmetry and triangle-inequality of the Hellinger
distance, we can simplify the local sensitivity.
\begin{lem}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}

(Proof is in Appendix IV.)

\subsubsection{Exponential Mechanism with Hellinger Metric and Smoothed Sensitivity ($\hexpmech$)}
\label{subsec_hexpmech}
In this section, we explore a new mechanism in Exponential family, which can achieve good accuracy and differential privacy simultaneously, by scaling the noise to a new sensitivity. We first define the new sensitivity we are going to use and then present the mechanism.
\begin{definition}
\label{def_gamma_smooth}
Given data set $\dataobs$ of size $n$, we define $\gamma -$smooth sensitivity of Bayesian inference process w.r.t. the Hellinger distance $S(\dataobs)$ is:
\begin{equation}
  \label{eq_smooth}
   S(\dataobs)=\max_{\dataobs'' \in \{0,1\}^{n}}\bigg \{ \frac{1}{\frac{1}{LS(\dataobs'')} +\gamma \cdot d(\dataobs,\dataobs'')} \bigg \},
\end{equation}
where $d$ is the Hamming distance between two data sets.
\end{definition}

\begin{thm}
\label{thm_gamma_smooth}
Given data set $\dataobs$, the $\gamma -$smooth sensitivity of Bayesian inference process w.r.t. the Hellinger distance $S(\dataobs)$ satisfying:%\\
whenever $\adj{\dataobs}{\dataobs'}$, 
%\begin{equation}
%\label{eq_gamma_smooth}
$\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
%\end{equation}
\end{thm}
(Proof is in Appendix I.)

>>>>

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
    \item A full proof of $(\epsilon, \delta)$-differential privacy of the mechanism is given in the paper. 
	\item We explored two ways to improve the accuracy: 1) calibrating noise to the sensitivity of a metric over distributions
      (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) Using a smooth upper bound on the local sensitivity and scale
      the noise to this smooth bound rather than global sensitivity, to improve the mechanism accuracy.
  \item We implemented the new proposed mechanism and other art-of-state mechanisms, comparing their performances in terms of accuracy and privacy.
\end{itemize}

@@

\noindent \textbf{Problem setting.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in Section \ref{sec_background}.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta and Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a private version of the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
However, distance between
distributions are more naturally  measured by distribution
metrics such as $f$-divergences, e.g. Hellinger distance, etc., than the $\ell_1$ norm shown above.
{\color{blue} still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

We will use the \emph{Hellinger} distance $\hellinger(\cdot,\cdot)$.
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the square of the Hellinger distance between them is defined as:
\[
\hellinger(P,Q)^2 =\frac{1}{2} \int \bigg(\frac{dP}{d \lambda} - \frac{dQ}{d \lambda}\bigg)^2 d \lambda
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$, we can unfold the above and derive: 
\[
 \hellinger(P,Q) =  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
Our choice to use Hellinger distance is motivated by three facts:
\begin{itemize}
  \item It simplifies calculations in the case of the probabilistic models considered here.

@@

  \item The accuracy can be improved by using a smooth bound on Hellinger distance's local sensitivity. Indeed, as shown in Figure \ref{fig_sensitivity}, there can be a gap
    between the global sensitivity and the local sensitivity.
\begin{figure}[ht]
\centering
\includegraphics[width=0.32\textwidth]{sensitivity_2d.eps}
\caption{Local sensitivity of $\hellinger$. Data size is 100. $\alpha$ is the number of 1s.}
\label{fig_sensitivity}
\end{figure}
\end{itemize}

\section{Private Mechanisms}
\label{sec_mechs}
\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
We start off by presenting a mechanism based on the Laplace primitive.
The mechanism, denoted by $\lapmech$, takes in input the data $\dataobs$, a mechanism, $\bysinfer$, that is used to perform inference on some parameter,
a prior and a likelihood function, and the $\epsilon$ parameter measuring the privacy level.
Now, $\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$,
simply adds Laplace noise to the parameters of the posterior distributions. The mechanisms also performs a
post-processing phase by flooring and  truncating the noisy parameters so that they belong to the right interval.
In the case of the Beta-Binomial model the output produced is: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$
with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into
$[0,n]$. This mechanism is described by the pesudocode in Figure \ref{mech:lapmech}.\\
Since we are adding noise directly to the parameters, the sensitivity should be calibrated w.r.t $\ell_1$-norm. That is,
the sensitivity is defined as:
\[
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
\]
This value can be quickly seen as being upper bounded by  the dimensionality. (e.g. 2 for the Beta-Binomial model and  $k$ for
the Dirichlet-Multinomial model).
  \begin{algorithm}
  \caption{$\lapmech$ for the Beta-Binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad  $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad  $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad  $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ \\
  % \RETURN $a_{k+1}$.
  \noindent {\bf return} $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{Improving the sensitivity ($\ilapmech$)}
 The accuracy of $\lapmech$ can be reduced to a constant.
 Indeed, two adjacent data sets $\dataobs, \dataobs'$  induce posterior distributions -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ --
 whose parameters differ by at most 2 w.r.t $\ell_1$ norm. This because we can see the parameters of the distributions as an histogram.
 Hence, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.
 {\color{red} Notice that we are using the bounded notion of differential privacy \cite{????}.}
 \todo{why 1 in 2 dimensions?}
 We denote the mechanism which calibrates the noise to this smaller sensitivity as $\ilapmech$.
 $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0) $
 where $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in the Beta-binomial model and
 $\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots,  \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}$
 with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Figure \ref{mech:ilapmech}. We will also from now on
 write $\ilapmech(\dataobs)$ when the other parameters are obvious from the context.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-Multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

\subsection{A new family of mechanisms}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{??} where the scoring function is based on a metric over
probability distributoins.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{??} proves privacy $\epsilon$ differential privacy of $\expmech$.

\subsubsection{A second approach based on local sensitivity: $\lexpmech$}
\label{subsec_emls}
Figure \ref{??} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
{\Pr}[z=r] = \frac
  {\exp(\frac{\epsilon u(x,r)}{2 LS(\dataobs)})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon u(x,r')}{2 LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix \ref{??}.

\subsubsection{A third approach based on smooth sensitivity: $\hexpmech$}
\label{subsec_hexpmech}
In this section, we explore a new mechanism in Exponential family, which can achieve good accuracy and $\epsilon$-differential privacy simultaneously,
by scaling the noise to a smooth upper bound on the local sensitivity.
We first define the smooth sensitivity on a input $\dataobs$. The following definition is specific to binary data, that is it will
be useful to privately release a posterior Beta distribution. The extension to categorical data is straightforward.
\begin{definition}
\label{def_gamma_smooth}
Given a data set $\dataobs$ of size $n$, we define the $\gamma$-smooth sensitivity of $\hellinger(\bysinfer(\dataobs),\cdot)$ as:
\begin{equation}
  \label{eq_smooth}
   S(\dataobs)=\max_{\dataobs' \in \{0,1\}^{n}}\bigg \{ \frac{1}{\frac{1}{LS(\dataobs')} +\gamma \cdot d(\dataobs,\dataobs')} \bigg \},
\end{equation}
where $d$ is the Hamming distance between two data sets.
\end{definition}

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\{0,1\}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
The interested reader will find the proof in Appendix \ref{??}.
We are now ready to define the mechanism $\hexpmech$. $\hexpmech(\dataobs)$ outputs a candidate $r \in \candidateset$ with probability
\[\frac {\exp\big(\frac{-\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)}\big)}
{\unomalizer{\dataobs}}.
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  $\hellinger(\bysinfer(\dataobs),\cdot)$. {\color{red}{why do we choose 1???}}

As already mentioned $\expmech, \lexpmech, \sexpmech$ extends straightforwardly to the Dirichlet-multinomial system.
Indeed, the Hellinger distance between two Dirichlet distribution, unfolds in:
\[
  \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}},
\]
and the $\candidateset$ will consist of the set of all possible posterior Dirichlet
distributions candidates. Also, the smooth sensitivity $S(\dataobs)$
in (\ref{thm_gamma_smooth}) will be computed by letting $\dataobs'$ range
over all the elements in $\datauni^{n}$ adjacent to $\dataobs$. Notice
that $\candidateset$ has $\binom{n + 1}{m - 1}$ elements in this case. {\color{red}{what is m here?}}
